---
title: "Learning activity - Bayesian Data Analysis"
author: "Eva Ascarza and Tian Zheng"
date: "December 21, 2015"
output: html_document
---

In this learning activity, we will first simulate some customer behavior data according to a model and then we will use Bayesian modeling and `rstan` to model the simulated data. Simulation is a great way of learning about how a method works.

If you have not done so already, you need to first install `rstan` by typing `install.packages("rstan"")` in your console windown (lower left panel in Rstudio).

```{r}
require(rstan)
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores()-2)
```

### Simulate the data
We will simulate a data set with 5000 invidiuals drawn from a population of musume members. Among museum members,

+ $X_1$ `gender`: Approximately 60% are females. 
+ $X_2$ `fotography`: 50% expressed interest for fotography.
+ $X_3$ `sculpture`:30% expressed interest for sculpture.
+ $X_4$ `digital_content`: The average number of digital contents consumed by the members is 0.5.
+ $X_5$ `months_since`: The average length since last visit is 5 months. 

```{r}
N <- 5000
# we assume 5000 members
female <- rbinom(N,1,0.6) 
# simulate 60% of female
fotography <- rbinom(N,1,0.5) 
# simulate 50% of customers who expressed interest for fotography
sculpture <- rbinom(N,1,0.3) 
# simulate 50% of customers who expressed interest for sculpture
digital_content <- rpois(N,0.5) 
# simulate digital content
months_since <- rpois(N,5) 
# simulate months since last visit, with a max of 12
months_since[months_since>12] <- 12
```

### simulate renewal behavior and prepare the data
We are interested in whether these members will renew their membership. $Y$ (`renewal`) equals 1 if the member decides to renew, 0 otherwise. We assume the following [probit model](https://en.wikipedia.org/wiki/Probit_model) for Y given X.

$$
  P(Y=1 | X_1, X_2, X_3, X_4, X_5) = \Phi(\beta_0 +\beta_1 X_1 +\beta_2 X_2 + \beta_3 X_3 +\beta_4 X_4 + \beta_5 X_5)
$$
where $\Phi(\cdot)$ is the [cumulative distribution function (CDF)](https://en.wikipedia.org/wiki/Cumulative_distribution_function) of the standard normal distribution. 

```{r}
beta0 <- 0.6
beta1 <- 0.9
beta2 <- 0.6
beta3 <- 0.01
beta4 <- -0.01
beta5 <- -0.2

prob_simul <- pnorm(beta0 + beta1*female + beta2*fotography + beta3*sculpture + beta4*digital_content + beta5*months_since)
renewal <- rbinom(N,1,prob_simul) 
```

### Prepare data for Stan
We first create an X matrix that combines all the input variables, including a column corresponding to the intercept. `cbind` is the `R` command for combining vectors as columns. 
```{r}
X <- cbind(rep(1,N),
           female,
           fotography,
           sculpture,
           digital_content,
           months_since)

# assign column names
colnames(X) <- c("intercept",
                 "female",
                 "fotography",
                 "sculpture",
                 "digital_content",
                 "months_since")
K<-dim(X)[2]
```

### Code for Stan

`stan` uses a specific modeling syntax. It requires the specification of the types of data and parameters, in addition to model statements. 

```{r}
probit <- '
data{
  int<lower=0> N; # number of observations
  int<lower=0> K; # number of parameters

  int<lower=0,upper=1> y[N];
  vector[K] X[N];
}
parameters{
  vector[K] beta;
}
model{
  beta ~ cauchy(0,5);

for(j in 1:N)
  y[j] ~ bernoulli(Phi_approx(dot_product(X[j],beta)));
}
'
```

### Run STAN

The initialization step of `stan` may take a little while but the running time should
be just a couple of minuets. 

```{r}
data_list <- list(N=N,K=K,y=renewal,X=X)
fit <- stan(model_code=probit,
            data=data_list,
            warmup=2000,
            iter=4000,
            chains=2)
```

### SUMMARY of restuls

In the summary of output, the posterior distribution for each model coefficient is summaried using its percentiles. 

```{r}
print(fit)
fitlist <- extract(fit)
```

#### Convergence

Convergence plots show whether the draws from posterior distributions are well mixed together. 

```{r}
par(mfrow=c(2,3))
plot(fitlist$beta[,1],type="l",xlab="",ylab="Intercept")
plot(fitlist$beta[,2],type="l",xlab="",ylab="Gender Coefficient")
plot(fitlist$beta[,3],type="l",xlab="",ylab="Fotography Coefficient")
plot(fitlist$beta[,4],type="l",xlab="",ylab="Sculpture Coefficient")
plot(fitlist$beta[,5],type="l",xlab="",ylab="Digital content Coefficient")
plot(fitlist$beta[,6],type="l",xlab="",ylab="Months since login")
```

#### Histograms of posterior distributions

Posterior distributions of model coefficients.

```{r}
par(mfrow=c(2,3))  
par(bg="white", fg="black", 
    col.lab="black", col.axis="black", 
    col.main="black", lwd=1)
#Gender
param <- fitlist$beta[,2]
min <- min(0,min(param))
max <- max(0,max(param))
hist(param,breaks=seq(min-0.05,max+0.05,0.05),main="Female",xlab="", ylab="",yaxt='n')
axis(1, lab=T , lwd=2)
abline(v=beta1, col=2)
#PHOTO
param <- fitlist$beta[,3]
min <- min(0,min(param))
max <- max(0,max(param))
hist(param,breaks=seq(min-0.04,max+0.04,0.04),main="Photography",xlab="", ylab="",yaxt='n')
axis(1, lab=T , lwd=2)
abline(v=beta2, col=2)
#SCULPTURE
param <- fitlist$beta[,4]
min <- min(0,min(param))
max <- max(0,max(param))
hist(param,breaks=seq(min-0.03,max+0.03,0.03),main="Sculpture",xlab="", ylab="",yaxt='n')
axis(1, lab=T , lwd=2)
abline(v=beta3, col=2)
#ONLINE CONTENT
param <- fitlist$beta[,5]
min <- min(0,min(param))
max <- max(0,max(param))
hist(param,breaks=seq(min-0.02,max+0.02,0.02),main="Downloads",xlab="", ylab="",yaxt='n')
axis(1, lab=T , lwd=2)
abline(v=beta4, col=2)
#VISIT
param <- fitlist$beta[,6]
min <- min(0,min(param))
max <- max(0,max(param))
hist(param,breaks=seq(min-0.01,max+0.01,0.01),main="Months since visit",xlab="", ylab="",yaxt='n')
axis(1, lab=T , lwd=2)
abline(v=beta5, col=2)
```
